---
title: "SpCas9 Validation Set"
output: html_document
author: "Peter DeWeirdt"
date: "November 7, 2018"
fontsize: 11pt
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(pheatmap)
library(gdata)
library(tidyr)
library(reshape2)
library(ggplot2)
library(plyr)
library(igraph)
library(RColorBrewer)
library(ggrepel)
library(gridExtra)
library(fossil)
library(e1071)
library(data.table)
setwd('~/Documents/csci-e-29/final_project-PeterDeWeirdt/data/raw/')

opts_chunk$set(echo = FALSE, warning = FALSE)
```

## Selecting Genes

We consider 47 genes tested across 3 drug assays: Selumetinib and A133 in Meljuso cells, as well as Talazoparib (low dose) in A375 cells. In order to select genes from these assays, we look for genes such that:

1. Pearson correlation between replicates is significant at the 1e-4 level
2. The Average of the Avg.LFC = (Rep.A.LFC + Rep.B.LFC)/2 for gene and condition specific guides is +/- one standard deviation from zero
3. The gene is not already in the training set

With this criteria we select nine genes visualized below:

```{r filt}
meljuso_raw = read.csv('/Users/pdeweird/Documents/csci-e-29/final_project-PeterDeWeirdt/data/raw/Meljuso_Cas9_Test_Filt.csv')
a375_raw = read.csv('/Users/pdeweird/Documents/csci-e-29/final_project-PeterDeWeirdt/data/raw/A375_Cas9_Test_Filt.csv')
a375_raw['Cell'] = 'A375'
a375_30mer = inner_join(a375_raw, meljuso_raw %>% select(c('Construct.Barcode', 'X30mer')) %>%
                          distinct(), by = 'Construct.Barcode') 
meljuso_annotation = inner_join(meljuso_raw, a375_raw %>% select(c('Construct.Barcode', 'Annotation')), 
                                by = 'Construct.Barcode')
joined_data = bind_rows(a375_30mer, meljuso_annotation) %>%
  dplyr::mutate(Avg.LFC = (Rep.A.LFC.From.Dropout + Rep.B.LFC.From.Dropout)/2) %>%
  dplyr::mutate(gene_condition = paste(Gene, Condition, sep = '_'))

conditions = unique(joined_data$Condition)
genes = unique(joined_data$Gene)
rows = length(conditions)*length(genes)
gene_summary = data.frame(cell = rep(NA, rows), condition = rep(NA, rows), gene = rep(NA, rows), 
                          Avg.Avg.LFC.pval = rep(NA, rows), 
                          Avg.Avg.LFC = rep(NA, rows), pearson = rep(NA, rows), pearson.pvalue = rep(NA, rows),
                          Avg.Avg.LFC.sd = rep(NA, rows))
i = 1
for (condition in conditions) {
  condition_df = joined_data %>%
    filter(Condition == condition)
  cond_sd = (condition_df %>% select(Avg.LFC) %>% unlist() %>% sd())
  for (gene in genes) {
    gene_df = condition_df %>%
      filter(Gene == gene)
    no_gene_df = condition_df %>%
      filter(Gene != gene)
    gene_summary[i, 'cell'] = unique(gene_df$Cell)
    gene_summary[i, 'condition'] = condition
    gene_summary[i, 'gene'] = gene
    avg.avg.lfc = mean(gene_df$Avg.LFC)
    gene_summary[i, 'Avg.Avg.LFC'] = avg.avg.lfc
    pearson = cor.test(gene_df$Rep.A.LFC.From.Dropout, gene_df$Rep.B.LFC.From.Dropout)
    gene_summary[i, 'pearson'] = pearson$estimate
    gene_summary[i, 'pearson.pvalue'] = pearson$p.value
    n_guides = nrow(gene_df)
    null_avg_avg_lfc = rep(NA, n_guides)
    n_resamps = 1000
    for(j in 1:n_resamps) {
      null_avg_avg_lfc[j] = mean(sample(condition_df$Avg.LFC, n_guides))
    }
    
    gene_summary[i, 'Avg.Avg.LFC.pval'] = min(sum(avg.avg.lfc < null_avg_avg_lfc), 
                                              sum(avg.avg.lfc > null_avg_avg_lfc))/n_resamps
    gene_summary[i, 'Avg.Avg.LFC.sd'] =  abs(avg.avg.lfc) > cond_sd                                 
    gene_summary[i, 'n_guides'] = n_guides
    i = i + 1
  }
}
train_genes = c('CCDC101', 'CUL3', 'HPRT1', 'MED12', 'NF1', 'NF2', 'TADA1', 'TADA2B')
test_gene_summary = gene_summary %>% filter(Avg.Avg.LFC.sd == TRUE & pearson.pvalue < 1e-4 & !(gene %in% train_genes)) %>%
  dplyr::mutate(gene_condition = paste(gene, condition, sep = '_'))
test_data = joined_data %>% filter(gene_condition %in% test_gene_summary$gene_condition)
ggplot(test_data)+
  aes(y=Rep.B.LFC.From.Dropout, x=Rep.A.LFC.From.Dropout, color = Condition) +
  geom_point(alpha = 0.5) +
  theme_classic() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline(slope = 1, intercept = 0, linetype = 'dotted') +
  facet_wrap('Gene') +
ggsave('/Users/pdeweird/Documents/csci-e-29/final_project-PeterDeWeirdt/data/figures/testdata.pdf')

```

From these genes we obtain a test set of 2517 guides. 

```{r output}
output_data = test_data %>%
  group_by(Gene) %>%
  dplyr::mutate(percentile = rank(abs(Avg.LFC))/n()) %>%
  write.csv('/Users/pdeweird/Documents/csci-e-29/final_project-PeterDeWeirdt/data/raw/SpCas9_test_guides.csv', 
             row.names = FALSE)
```


